---
title: "BrainChart calibration - assessing accuracy of centile method"
format: html
---

```{r, echo=FALSE}
knitr::read_chunk("calibration_example_hbn_accuracy.R")
```

# Introduction

This tutorial demonstrates how to assess the accuracy of the proposed centile-informed calibration compared to the original Braincharts' method on small samples on the HBN dataset. The first thing to do is source the script that loads and performs the original calibration:

```{r Setup, message = FALSE, warning = FALSE}
BrainChartFolder <- dirname(normalizePath("~/Projects/brainchartscentiles/calibrate_braincharts_centiles.R"))
source(file.path(BrainChartFolder, 'calibration_example_hbn.R'))
```

## Generate synthetic timepoint

For each site, we generate a synthetic repeated-measures sample, approximately one year post-baseline. The following code inserts the predicted distribution parameters into the original dataframe, copies the data to a *fake* dataframe with altered parameters. The new cortical thickness values are then generated and the study names are updated to be different to the original:

```{r Setup, message = FALSE, warning = FALSE}
brainChartsDF$mu.wre <- NA
brainChartsDF$sigma.wre <- NA
brainChartsDF$nu.wre <- NA

brainChartsDF[row.names(fullSampleCBIC$DATA.PRED2), 'mu.wre'] <- fullSampleCBIC$DATA.PRED2$mu.wre
brainChartsDF[row.names(fullSampleCBIC$DATA.PRED2), 'sigma.wre'] <- fullSampleCBIC$DATA.PRED2$sigma.wre
brainChartsDF[row.names(fullSampleCBIC$DATA.PRED2), 'nu.wre'] <- fullSampleCBIC$DATA.PRED2$nu.wre

brainChartsDF[row.names(fullSampleCUNY$DATA.PRED2), 'mu.wre'] <- fullSampleCUNY$DATA.PRED2$mu.wre
brainChartsDF[row.names(fullSampleCUNY$DATA.PRED2), 'sigma.wre'] <- fullSampleCUNY$DATA.PRED2$sigma.wre
brainChartsDF[row.names(fullSampleCUNY$DATA.PRED2), 'nu.wre'] <- fullSampleCUNY$DATA.PRED2$nu.wre

brainChartsDF[row.names(fullSampleRU$DATA.PRED2), 'mu.wre'] <- fullSampleRU$DATA.PRED2$mu.wre
brainChartsDF[row.names(fullSampleRU$DATA.PRED2), 'sigma.wre'] <- fullSampleRU$DATA.PRED2$sigma.wre
brainChartsDF[row.names(fullSampleRU$DATA.PRED2), 'nu.wre'] <- fullSampleRU$DATA.PRED2$nu.wre

brainChartsDF[row.names(fullSampleSI$DATA.PRED2), 'mu.wre'] <- fullSampleSI$DATA.PRED2$mu.wre
brainChartsDF[row.names(fullSampleSI$DATA.PRED2), 'sigma.wre'] <- fullSampleSI$DATA.PRED2$sigma.wre
brainChartsDF[row.names(fullSampleSI$DATA.PRED2), 'nu.wre'] <- fullSampleSI$DATA.PRED2$nu.wre

# generate a new table with altered distribution parameters
brainChartsDFFake <- brainChartsDF

# add timepoint effects
muNudge <- 0.1
sigmaNudge <- 0.1
brainChartsDFFake$mu.wre <- brainChartsDF$mu.wre + muNudge
brainChartsDFFake$sigma.wre <- brainChartsDF$sigma.wre + sigmaNudge

# add a year to the ages and some randomness
brainChartsDFFake$age_days <- brainChartsDFFake$age_days + 365.25 + abs(rnorm(nrow(brainChartsDFFake), mean = 0, sd = 10))
# generate variables from altered distributions
brainChartsDFFake$CT <- rGGalt(nrow(brainChartsDFFake), exp(brainChartsDFFake$mu.wre), exp(brainChartsDFFake$sigma.wre), brainChartsDFFake$nu.wre) * 10000

brainChartsDFFake$study <- paste0(brainChartsDFFake$study, "2")
```

The dataframe *brainChartsDFFake* contains the repeated measures timepoint. We now subsample this frame to create small samples:

```{r CreateSmallSamples, message = FALSE, warning = FALSE}
n <- 50
subFakeCBIC <- filter(brainChartsDFFake, study == "HBNCBIC2")
P <- randperm(1:nrow(subFakeCBIC))
subFakeCBIC <- subFakeCBIC[P[1:n],]
subFakeCUNY <- filter(brainChartsDFFake, study == "HBNCUNY2")
P <- randperm(1:nrow(subFakeCUNY))
subFakeCUNY <- subFakeCUNY[P[1:n],]
subFakeRU <- filter(brainChartsDFFake, study == "HBNRU2")
P <- randperm(1:nrow(subFakeRU))
subFakeRU <- subFakeRU[P[1:n],]
subFakeSI <- filter(brainChartsDFFake, study == "HBNSI2")
P <- randperm(1:nrow(subFakeSI))
subFakeSI <- subFakeSI[P[1:n],]
```

This code selects 50 random subjects from each sample.

## Calibration - Phase 1

The normal calibration is now performed on the small sample using the original Braincharts' method:

```{r Calib1, message = FALSE, warning = FALSE}
subFakeCBICOut <- calibrateBrainCharts(subFakeCBIC, phenotype = "CT")
subFakeCUNYOut <- calibrateBrainCharts(subFakeCUNY, phenotype = "CT")
subFakeRUOut <- calibrateBrainCharts(subFakeRU, phenotype = "CT")
subFakeSIOut <- calibrateBrainCharts(subFakeSI, phenotype = "CT")
```


## Calibration - Phase 2

We now run the calibration on the small samples using centile-informed calibration method using the large sample data:

```{r Calib2, message = FALSE, warning = FALSE}
subFakeCBICOutQuantile <- calibrateBrainChartsIDQuantilePenalty(subFakeCBIC, phenotype = "CT", largeSiteOutput = fullSampleCBIC)
subFakeCUNYOutQuantile <- calibrateBrainChartsIDQuantilePenalty(subFakeCUNY, phenotype = "CT", largeSiteOutput = fullSampleCUNY)
subFakeSIOutQuantile <- calibrateBrainChartsIDQuantilePenalty(subFakeSI, phenotype = "CT", largeSiteOutput = fullSampleSI)
subFakeRUOutQuantile <- calibrateBrainChartsIDQuantilePenalty(subFakeRU, phenotype = "CT", largeSiteOutput = fullSampleRU)
```

## Calibration of full fake sample

We now calibrate the entire fake sample. The computed parameters are used as ground truth for evaluation:

```{r Calib2, message = FALSE, warning = FALSE}
fakeSampleCBIC <- calibrateBrainCharts(filter(brainChartsDFFake, study == "HBNCBIC2"), phenotype = phenotype)
fakeSampleCUNY <- calibrateBrainCharts(filter(brainChartsDFFake, study == "HBNCUNY2"), phenotype = phenotype)
fakeSampleSI <- calibrateBrainCharts(filter(brainChartsDFFake, study == "HBNSI2"), phenotype = phenotype)
fakeSampleRU <- calibrateBrainCharts(filter(brainChartsDFFake, study == "HBNRU2"), phenotype = phenotype)
```


## Make plots for evaluation

```{r MakePlots, message = FALSE, warning = FALSE}
fullSampleCBIC$DATA.PRED2$fitting <- 'large sample (cross-sectional method)'
subFakeCBICOutQuantile$DATA.PRED2$fitting <- 'small site (centile method)'
subFakeCBICOut$DATA.PRED2$fitting <- 'small site (cross-sectional method)'
fakeSampleCBIC$DATA.PRED2$fitting <- 'small site (ground truth)'

fullSampleCBIC$DATA.PRED2$sample <- "large"
fullSampleCBIC$DATA.PRED2
subFakeCBICOutQuantile$DATA.PRED2$sample <- "small"
subFakeCBICOut$DATA.PRED2$sample <- "small"
fakeSampleCBIC$DATA.PRED2$sample <- "small"

T <- bind_rows(fullSampleCBIC$DATA.PRED2, subFakeCBICOutQuantile$DATA.PRED2, subFakeCBICOut$DATA.PRED2, fakeSampleCBIC$DATA.PRED2)

#mapping <- c(V1 = "large", V5 = "small")
#T$sample <- mapping[T$study]

# dont need to plot the ground truth CT measures for the small site
T$meanCT2Transformed[T$fitting == "small site (ground truth)"] <- NA

# meanCT2Transformed is the transformed cortical thickness
# PRED.l250.wre, PRED.m500.wre, PRED.u750.wre are the estimated 25th, 50th, and 97.5th Quantiles after site-effect corrections

ggplot(T, aes(x = age_days)) +
  geom_point(aes(y = meanCT2Transformed * 10000, shape = sample), alpha = 0.5) +
  geom_line(aes(y = PRED.l250.wre * 10000, colour = fitting)) + 
  geom_line(aes(y = PRED.m500.wre * 10000, colour = fitting), linewidth = 1) +
  geom_line(aes(y = PRED.u750.wre * 10000, colour = fitting)) +
  labs(x = "Age (years)", y = "CT", colour = "Estimation Method")
```