---
title: "Small sample, repeated measure, calibration with BrainChart"
format: html
---

# Introduction

This site contains two tutorials explaining calibration using the method published
in:

# Installation

This software runs as as addon to the Braincharts software, which was written in R. In a shell, obtain the software from GitHub as follows:

```{bash eval=FALSE}
git clone https://github.com/NCHADataPlatform/brainchartscentiles.git
cd brainchartscentiles
git submodule init .
git submodule update
```

# R dependencies

In R, install the required packages for Braincharts:

```{r eval=FALSE}
install.packages(c('gamlss', 'tidyverse'))
```

Then install the dependencies for the example scripts on the github repository:

```{r eval=FALSE}
install.packages(c('ggplot2', 'pracma'))
```

The previous two commands only need to be executed once. 

For each R session, execute the following to import the software:

```{r eval=FALSE}
source('<install directory>/calibrate_braincharts_centiles.R')
```

# Tutorial 1:

This is where you should start after installing the software

[Calibration tutorial](calibration_example.qmd)

# Tutorial 2:

This is minimal example of the procedures used to assess calibration accuracy
as reported in the paper.

[Accuracy assessment](calibration_accuracy.qmd)

